Cross-validation of regression models: mixed model examples (edited)

### 2023-10-30 scratch/notes_georges/gm_addition_to_cv-edited.Rmd
move __notes_georges to scratch

```{r plot1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Hierarchical data set with 10 clusters"
plot <- xyplot(y ~ x, data=Data[ 1:50,], group=group,
               ylim=c(0, 20),
               par.settings=list(superpose.symbol=list(pch=1, cex=0.7))) +
    layer(panel.ellipse(..., center.cex=0))
plot
```

@@

Data_long <- reshape(Data[1:50, ], direction = "long", sep = ".", 
              timevar = "effect", varying = grep("\\.", names(Data.10)))
Data_long$id <- 1:nrow(Data_long)
Data_long <- reshape(Data_long, direction = "long", sep = "_", 
              timevar = "modelcode",  varying = grep("_", names(Data_long)))

@@

```{r plot-fits-mod0}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x as a predictor"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim=c(-20, 40),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod2}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x and the group mean of x as predictors"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" &

                           effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim = c(0, 20),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```

>>>>

```{r plot1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Hierarchical data set with 10 clusters"
plot <- xyplot(y ~ x, data=Data[ 1:50,], group=group,
               ylim=c(4, 16),
               par.settings=list(superpose.symbol=list(pch=1, cex=0.7))) +
    layer(panel.ellipse(..., center.cex=0))
plot
```

@@

Data_long <- reshape(Data[1:50, ], direction = "long", sep = ".", 
              timevar = "effect", varying = grep("\\.", names(Data[1:50, ])))
Data_long$id <- 1:nrow(Data_long)
Data_long <- reshape(Data_long, direction = "long", sep = "_", 
              timevar = "modelcode",  varying = grep("_", names(Data_long)))

@@

```{r plot-fits-mod0}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x as a predictor"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim=c(4, 16),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod2}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x and the group mean of x as predictors"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" &

                           effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim=c(4, 16),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```

### 2023-10-30 scratch/notes_georges/gm_addition_to_cv-edited_v2_tmp.Rmd
move __notes_georges to scratch

# Parameters:
set.seed(9693) # set.seed(123) 
Nb <- 100     # number of groups
Nw <- 5      # number of individuals within groups
Bb <- 0      # between-group regression coefficient on group mean
SDre <- 2.0  # between-group SD of random level relative to group mean of x
SDwithin <- 0.5  # within group SD
Bw <- 1  # within group effect of x

Data <- data.frame(
  group = factor(rep(1:Nb, each=Nw)),
  x = 20 + seq(1, Nb*Nw)
) |>
  within(
    {
      xm  <- ave(x, group, FUN = mean) # within-group mean
      y <- 10 +
        Bb * xm +                  # contextual effect
        Bw * (x - xm) +            # within-group effect
        SDre * rnorm(Nb)[group] +  # random level by group
        SDwithin * rnorm(Nb*Nw)    # random error within groups
    }
  )
```

@@

```{r plot-fits-mod0}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x as a predictor"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim=c(4, 16),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod2}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x and the group mean of x as predictors"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" &

@@

xyplot(CV ~ model, data=dout, groups=cv_method,
       par.settings =
         list(superpose.line=list(lty=c(2, 3), lwd=3), 
              superpose.symbol=list(pch=15:16, 
                                    cex=1.5)),
       type="b",
       auto.key=TRUE,
       scales=list(y=list(log=TRUE)))
```

>>>>

# Parameters:
set.seed(9693) 
Nb <- 100     # number of groups
Nw <- 5      # number of individuals within groups
Bb <- 0      # between-group regression coefficient on group mean
SDre <- 2.0  # between-group SD of random level relative to group mean of x
SDwithin <- 0.5  # within group SD
Bw <- 1  # within group effect of x

Data <- data.frame(
  group = factor(rep(1:Nb, each=Nw)),
  x = 20 + rep(seq(1, 50), length.out = Nw*Nb)
) |>
  within(
    {
      xm  <- ave(x, group, FUN = mean) # within-group mean
      y <- 10 +
        Bb * xm +                  # contextual effect
        Bw * (x - xm) +            # within-group effect
        SDre * rnorm(Nb)[group] +  # random level by group
        SDwithin * rnorm(Nb*Nw)    # random error within groups
    }
  )
```

@@

```{r plot-fits-mod0}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & 
                         effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x as a predictor"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "fe"), 
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & 
                         effect == "re"), 
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  ylim=c(-15, 35),
  key=list(
    space="right",
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod2}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values with random intercept model and x and the group mean of x as predictors"
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" &

@@

xyplot(CV ~ model, data=dout, groups=cv_method,
       par.settings =
         list(superpose.line=list(lty=c(2, 3), lwd=3), 
              superpose.symbol=list(pch=15:16, 
                                    cex=1.5)),
       type="b",
       auto.key=list(reverse.rows=TRUE),
       scales=list(y=list(log=TRUE)))
```

### 2023-10-30 scratch/notes_georges/test.Rmd
move __notes_georges to scratch

# Parameters:
set.seed(9693) # set.seed(123) 
Nb <- 100     # number of groups
Nw <- 5      # number of individuals within groups
Bb <- 0      # between-group regression coefficient on group mean
SDre <- 2.0  # between-group SD of random level relative to group mean of x
SDwithin <- 0.5  # within group SD
Bw <- 1  # within group effect of x

@@

```{r plot1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Hierarchical data set with 10 clusters"
plot <- xyplot(y ~ x, data=Data[ 1:50,], group=group,
               ylim=c(0, 20),
               par.settings=list(superpose.symbol=list(pch=1, cex=0.7))) +
    layer(panel.ellipse(..., center.cex=0))
plot
```

@@

Data_long <- reshape(Data[1:50, ], direction = "long", sep = ".", 
              timevar = "effect", varying = grep("\\.", names(Data.10)))
Data_long$id <- 1:nrow(Data_long)
Data_long <- reshape(Data_long, direction = "long", sep = "_", 
              timevar = "modelcode",  varying = grep("_", names(Data_long)))

@@

```{r cross-validation}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Cross-validation comparing random intercept models with varying fixed effects and with case- and cluster-based resampling."
modlist <- models("~ 1"=fit0, "~ x"=fit1, "~ x + xm"=fit2)
cvs_cases <- cv(modlist, data=Data, seed=9693)
cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)

@@

xyplot(CV ~ model, data=dout, groups=cv_method,
       par.settings =
         list(superpose.line=list(lty=c(2, 3), lwd=3), 
              superpose.symbol=list(pch=15:16, 
                                    cex=1.5)),
       type="b",
       auto.key=TRUE,
       scales=list(y=list(log=TRUE)))
```

>>>>

# Parameters:
set.seed(9665) # set.seed(123) # set.seed(9693)
Nb <- 10     # number of groups
Nw <- 5      # number of individuals within groups
Bb <- 0      # between-group regression coefficient on group mean
SDre <- 2.0  # between-group SD of random level relative to group mean of x
SDwithin <- 0.5  # within group SD
Bw <- 1  # within group effect of x

@@

```{r plot1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Hierarchical data set with 10 clusters"
plot <- xyplot(y ~ x, data=Data, group=group,
               ylim=c(0, 20),
               par.settings=list(superpose.symbol=list(pch=1, cex=0.7))) +
    layer(panel.ellipse(..., center.cex=0))
plot
```

@@

Data_long <- reshape(Data, direction = "long", sep = ".", 
              timevar = "effect", varying = grep("\\.", names(Data)))
Data_long$id <- 1:nrow(Data_long)
Data_long <- reshape(Data_long, direction = "long", sep = "_", 
              timevar = "modelcode",  varying = grep("_", names(Data_long)))

@@

```{r cross-validation}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Cross-validation comparing random intercept models with varying fixed effects and with case- and cluster-based resampling."

modlist <- models("~ 1"=fit0, "~ x"=fit1, "~ x + xm"=fit2)
cvs_cases <- cv(modlist, data=Data, seed=9693)
cvs_clusters <- cv(modlist, data=Data, cluster="group", seed=4576)

{

  # play
  cv0 <- replicate(10,cv(fit0))
  cv1 <- replicate(10,cv(fit1))
  cv2 <- replicate(10,cv(fit2))

  cv0
  cv1
  cv2

  cv12 <- cv(fit1)
  cv13 <- cv(fit1)
  cv14 <- cv(fit1)
  cv01 <- cv(fit0)
  cv12 <- cv(fit0)
  cv13 <- cv(fit0)
  cv14 <- cv(fit0)
  # 
  # 
  cv11
  cv12
  cv13
  cv14

  cvs_cases <- cv(modlist, data=Data)
  cvs_clusters <- cv(modlist, data=Data, cluster="group")

@@

  xyplot(CV ~ model, data=dout, groups=cv_method,
         par.settings =
           list(superpose.line=list(lty=c(2, 3), lwd=3), 
                superpose.symbol=list(pch=15:16, 
                                      cex=1.5)),
         type="b",
         auto.key=TRUE,
         scales=list(y=list(log=TRUE)))
}
```
